---
title: "Comparison with PTW"
output:
  "BiocStyle::html_document":
    dev: png
  "BiocStyle::pdf_document":
    latex_engine: lualatex
    df_print: "kable"
    dev: png
package: pow
author: "pow authors"
date: "`r format(Sys.Date(), '%F')`"
abstract: >
  A comparison with the technique PTW
vignette: >
  %\VignetteIndexEntry{Comparison_PTW}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pow)
library(ggplot2)
library(BiocParallel)
#library(GCIMS)
register(SnowParam(workers = 2, 
                   progressbar = interactive() && 
                     is.null(getOption("knitr.in.progress")), 
                   exportglobals = FALSE), 
         default = TRUE)
```

This vingette compares the PTW algorithm proposed by Eilers (2004), with the POW algorithm proposed by Férnandez (N.D.)
```{r create synthetic dataset}
set.seed(2424)
data <- synthetic_dataset(n_samples = 10,
                          n_peaks = 10,
                          length_out = 1500,
                          mov_peaks = 5,
                          intensity = 500,
                          random_intensity = 50)
```

```{r use samples from zenodo}
# samples_directory <- "ketones"
# dir.create(samples_directory, recursive = TRUE, showWarnings = FALSE)
# url <- "https://zenodo.org/record/7941230/files/AnalyticalStandards.zip?download=1"
# tmp_zipfile <- tempfile(fileext=".zip")
# curl::curl_download(url, tmp_zipfile)
# utils::unzip(
#   tmp_zipfile,
#   junkpaths = TRUE, exdir = samples_directory
# )
# annotations <- create_annotations_table(samples_directory)
```

```{r}
# samples_directory<- "/storage/projects/TargetML/KETONES"
# annotations <- read.csv(file.path(samples_directory,"annotations.csv"))
# annotations <- annotations[-c(26,27,32,34,36,37,38),]
```


```{r load dataset with GCIMS}
# dataset <- GCIMSDataset$new(
#   annotations,
#   base_dir = samples_directory,
#   on_ram = FALSE 
# )
# filterRt(dataset, rt = c(0, 1100)) # in s
# filterDt(dataset, dt = c(5, 16)) # in ms
# dataset1 <- smooth(dataset, rt_length_s = 3, dt_length_ms = 0.14)
# decimate(dataset, rt_factor = 1, dt_factor = 2)
# align_ip(dataset)
# dataset$realize()
# data <- dataset$getRIC()
```

For the POW algorithm we have to optimize the 'lambda2' and select the best reference sample:
```{r optimization of parameters}
#First we select the best reference
idx_y <- select_reference(data)
#We separate the dataset in the reference sample and the samples to align
X <- as.matrix(data[-idx_y, ])
y <- as.matrix(data[idx_y, ])
#Then we optimize the lambda_2 
n_samples <- nrow(X)
m <- length(y)
v <- rep(1, m)
iv <- seq(2, m - 1, by = 10)
v[iv] <- 0
W <- Matrix::Diagonal(x = v)
lambdas <- pracma::logspace(-2, 4, 31)
val<-compute_val_error(X, y, W, iv, lambdas)
opar = optimize_params(n_samples, lambdas, val$ti_ix, val$e_ix)
best_lambdas2 <- opar$best_params
```

Once we have the best reference and the best lambda_2 for each sample we apply the algorithm
```{r Apply}
P <- apply_pow(X, best_lambdas2, y, return_warps = TRUE)
XW <- P$samples
warps <- P$w
```
To use PTW you have to optimize the order of the polynomials
```{r function for optimization of PTW}
optimize_polynomial_order <- function(ric_sample, ric_ref) {
  correction_type_options <- seq.int(0, 10)
  poly_orders <- seq.int(1, 10)
  xi <- seq_len(length(ric_sample))
  corr <- numeric(length(poly_orders) + 1L)
  corr[1] <- stats::cor(ric_ref, ric_sample, use = "complete.obs")
  for (j in seq_along(poly_orders)) {
    poly <- rep(0, poly_orders[j])
    poly[2] <- 1
    corr[j + 1L] <- stats::cor(
      ric_ref,
      as.numeric(ptw::ptw(ref = ric_ref, samp = ric_sample, init.coef = poly)$warped.sample[1,]),
      use = "complete.obs"
    )
  }
  correction_type_options[which.max(corr)]
}
```

Then we apply the PTW algorithm:
```{r}
xp<-XW*0
wp<-XW*0
order_p<-c()
for (i in 1:(nrow(xp))){
  poly_order <- optimize_polynomial_order(X[i,], as.numeric(y))
  poly <- rep(0, poly_order + 1L)
  poly[2] <- 1
  align_result <- ptw::ptw(ref = as.numeric(y), samp = X[i,], init.coef = poly)
  xp[i,]<-align_result$warped.sample
  wp[i,]<-align_result$warp.fun
  order_p <- c(order_p,poly_order)
}
```

In here we can see the the comparison, first we see the raw samples, then the samples aligned using the POW algorithm and finally with the PTW alignment
```{r}
df_raw <- data.frame(t(X))
df_raw <- cbind(df_raw ,Time = c(1:nrow(df_raw)))
df_raw <- reshape2::melt(df_raw, "Time")
raw_data <- ggplot(data = df_raw, aes(x = Time, y = value, col = variable)) +
  geom_line() +
  ylab("Intensity") +
  guides(colour = "none")

df_pow <- data.frame(t(XW))
df_pow <- cbind(df_pow,Time = c(1:nrow(df_pow)))
df_pow <- reshape2::melt(df_pow, "Time")
pow_data <- ggplot(data = df_pow, aes(x = Time, y = value, col = variable)) + 
  geom_line() +
  ylab("Intensity") +
  guides(colour = "none")



df_ptw <- data.frame(t(xp))
df_ptw <- cbind(df_ptw, Time = c(1:nrow(df_ptw)))
df_ptw <- reshape2::melt(df_ptw, "Time")
ptw_data <- ggplot(data = df_ptw, aes(x = Time, y = value, col = variable)) +
  geom_line() +
  ylab("Intensity") +
  guides(colour="none")

cowplot::plot_grid(raw_data,pow_data,ptw_data)
```
```{r}
image(c(1:m), 
      c(1:nrow(X)), 
      t(X), 
      col = hcl.colors(10, "Grays"),  
      xlab = "Time", 
      ylab = "Sample") + 
  title("Image of Raw Samples")
```

```{r}
image(c(1:m), 
      c(1:nrow(XW)), 
      t(XW), 
      col = hcl.colors(10, "Grays"),  
      xlab = "Time", 
      ylab = "Sample") + 
  title("Image of aligned Spectra with POW")
```

```{r}
image(c(1:m), 
      c(1:nrow(xp)), 
      t(xp), 
      col = hcl.colors(10, "Grays"),  
      xlab = "Time", 
      ylab = "Sample") + 
  title("Image of aligned Spectra with PTW")
```

```{r}
dwp <- diff(wp)
df_wp <- data.frame(t(dwp))
df_wp <- cbind(df_wp,T=c(1:nrow(df_wp)))
df_wp <- reshape2::melt(df_wp,"T")
ggplot(data=df_wp, aes(x=T, y=value, col=variable))+geom_line()
```
